---
title: "Credit Card Profittability"
output: github_document
---

## Predicting weather a Credit Card account is profitable or not.

```{r setup, include=FALSE}

library(readr)
library(tree)
library(ROCR)
library(class)
library(ggplot2)

#Loading the data
creditDF <- read_csv("Data/Credit_Dataset.csv")

class_performance <- function(confusion_matrix){

  TP <- confusion_matrix[2,2]
  TN <- confusion_matrix[1,1]
  FP <- confusion_matrix[1,2]
  FN <- confusion_matrix[2,1]
  
  ##accuracy = total number of correct classifications/total number of classifications
  acc <- (TP+TN)/(TP+TN+FP+FN)
  
  ##TPR = Percent of actual positives identified as such (sensitivity)
  tpr <- TP/(TP+FN)
  
  ##TNR = Percent of actual negatives identified as such (specificity)
  tnr <- TN/(TN+FP)
  
  ##I'll leave it as an exercise for you to compute the other basic confusion matrix metrics
  
  ##return the list of metrics you want
  return(c(acc, tpr, tnr))
}


confusion_matrix <- function(preds, actuals, cutoff){
  
  classifications <- ifelse(preds>cutoff,1,0)
  
  ##careful with positives and negatives here!
  confusion_matrix <- table(actuals,classifications)
}



predict_and_classify <- function(treename, pred_data, actuals, cutoff){
  probs <- predict(treename, newdata = pred_data)[,2]
  classifications <- ifelse(probs>cutoff,1,0)
  accuracy <- sum(ifelse(classifications==actuals,1,0))/length(actuals)
  return(accuracy)
}


```

## Part 1: Data Preparation

1) Wrangling

```{r}
# Adding the profitable column
creditDF$PROFITABLE <- ifelse(creditDF$PROFIT >= 0 , 1, 0)

#Making CHK_ACCT, SAV_ACCT, HISTORY, JOB, and TYPE as factors
creditDF$CHK_ACCT <- as.factor(creditDF$CHK_ACCT)
creditDF$SAV_ACCT <- as.factor(creditDF$SAV_ACCT)
creditDF$HISTORY <- as.factor(creditDF$HISTORY)
creditDF$JOB <- as.factor(creditDF$JOB)
creditDF$TYPE <- as.factor(creditDF$TYPE)
creditDF$PROFITABLE <- as.factor(creditDF$PROFITABLE)
```

2) Setting the Seed and Splitting

```{r}

# Setting the seed
set.seed(12345)

# 30% as test data and 70% as rest
test_instn = sample(nrow(creditDF), 0.3*nrow(creditDF))
creditTestDF <- creditDF[test_instn,]
creditRestDF <- creditDF[-test_instn,]

# 25% as validation data and 75% as training data
valid_instn = sample(nrow(creditRestDF), 0.25*nrow(creditRestDF))
creditValidDF <- creditRestDF[valid_instn,]
creditTrainDF <- creditRestDF[-valid_instn,]

```


## Part 2: Logistic regression

1) Training a logistic regression model to predict your newly created categorical dependent variable PROFITABLE, using AGE, DURATION, RENT, TELEPHONE, FOREIGN, and the factors you created from CHK_ACCT, SAV_ACCT, HISTORY, JOB, and TYPE.

```{r}

#LogisticRegression
LogisticsRegression <- glm(PROFITABLE~AGE+DURATION+RENT+
                             TELEPHONE+FOREIGN+CHK_ACCT+
                             SAV_ACCT+HISTORY+JOB+TYPE, data = creditTrainDF, family = 'binomial')

#Predicting using the predict function
logisticPrediction <- predict(LogisticsRegression, newdata = creditValidDF, type = 'response')

logisticTestPrediction <- predict(LogisticsRegression, newdata = creditTestDF, type = 'response')


```


2) Plot the accuracy, sensitivity(TPR), and specificity(TNR) against all cutoff values (using the ROCR package) for the validation data. What is that maximum accuracy value? At what value of the cutoff is the accuracy maximized? 

```{r}

logisticPred <- prediction(logisticPrediction,creditValidDF$PROFITABLE)
logisticTPR = performance(logisticPred, measure = 'tpr')
logisticTNR <- performance(logisticPred, measure = 'tnr')
logisticACC <- performance(logisticPred, measure = 'acc')
logisticROC <- performance(logisticPred, measure = 'tpr', x.measure = 'fpr')

plot(logisticTPR, ylim=c(0,1))
plot(logisticTNR, add= T, col = 'red')
plot(logisticACC, add = T, col = 'blue')

best = which.max(slot(logisticACC,"y.values")[[1]])
maxACC = slot(logisticACC,"y.values")[[1]][best]

```

Maximum accuracy = 0.7542
Cutoff = 0.6619

3) ROC curves for both the training and validation data on the same chart.

```{r}

logisticPredictionTraining <- predict(LogisticsRegression, newdata = creditTrainDF, type = 'response')

logisticTrainPred <- prediction(logisticPredictionTraining, creditTrainDF$PROFITABLE)
logisticTrainROC = performance(logisticTrainPred, measure = "tpr", x.measure = "fpr")
plot(logisticTrainROC, col = 'red')
plot(logisticROC, add = T, col = 'blue')

```

Nothing Unexpected but the training data is expected to have a higher AUC

4) Plot the lift curve for the validation data

```{r}
liftValid = performance(logisticPred, measure = "lift", x.measure = "rpp")
plot(liftValid)
```

Maximum lift - 1.5625
For 20% of the loans lift is - 1.39
Lift of 1.3 maximum Positive predictions - 48%

## Part 3: Classfication Trees

1) Classification Tree algorithm to predict PROFITABLE using the training data and the variables you used in your logistic regression model. Experimenting with different tree sizes by modifying the number of terminal nodes in the tree. Using 10 values: 2, 4, 6, 8, 10, 12, 14, 16, 18, 20, as well as the full (unpruned) tree.

```{r}
creditTree = tree(PROFITABLE~AGE+DURATION+RENT+
                             TELEPHONE+FOREIGN+CHK_ACCT+
                             SAV_ACCT+HISTORY+JOB+TYPE, data = creditTrainDF)

fullTree0ACC = predict_and_classify(creditTree, creditValidDF, 
                                   creditValidDF$PROFITABLE, 0.5)

trainAcc = c()
validAcc = c()
tNodes = c()

for(i in c(2, 4, 6, 8, 10, 12, 14, 16, 18, 20)){
  prunedTree = prune.tree(creditTree, best = i)
  validAcc = c(validAcc, predict_and_classify(prunedTree, creditValidDF, 
                                   creditValidDF$PROFITABLE, 0.5))
  trainAcc = c(trainAcc, predict_and_classify(prunedTree, creditTrainDF, 
                                   creditTrainDF$PROFITABLE, 0.5))
  tNodes = c(tNodes, i)
}
treeDF = data.frame(tNodes, trainAcc, validAcc)
treeDF
```

2) Cutoff of 0.5 to classify accounts and measure the accuracy in the training and validation data for each tree. Plot the tree size versus accuracy in the training and validation data (respectively) and select the best tree size.

```{r}
ggplot() + 
  geom_line(data = treeDF, aes(x = tNodes, y = trainAcc), color = 'red') +
  geom_line(data = treeDF, aes(x = tNodes, y = validAcc), color = 'blue')
```

3) Plot the tree that results in the best accuracy in the validation data.

```{r}
prunedTree = prune.tree(creditTree, best = 6)
plot(prunedTree)
text(prunedTree,pretty=1)
```

The best accuracy is for 6 terminal nodes.

## Part4: KNN Algorithm

1) kNN algorithm for classification on the training data using the following variables: AGE, DURATION, RENT, TELEPHONE, FOREIGN, CHK_ACCT, SAV_ACCT, HISTORY, JOB, and TYPE

1) Trying ten values of k: 1, 3, 5, 7, 11, 15, 21, 25, 31, and 35.

```{r}
creditKNNDF <- read_csv("Data/Credit_Dataset.csv")
creditKNNDF$PROFITABLE <- ifelse(creditKNNDF$PROFIT >= 0 , 1, 0)
creditKNNDF$PROFITABLE <- as.numeric(creditKNNDF$PROFITABLE)
set.seed(12345)

# 30% as test data and 70% as rest
#creditKNNDF <- creditKNNDF[,c(2,3,4,6,7,10,12,17,18,20,24)]
test_instn = sample(nrow(creditKNNDF), 0.3*nrow(creditKNNDF))
creditKNNTestDF <- creditKNNDF[test_instn,]
creditKNNRestDF <- creditKNNDF[-test_instn,]

# 25% as validation data and 75% as training data
valid_instn = sample(nrow(creditKNNRestDF), 0.25*nrow(creditKNNRestDF))
creditKNNValidDF <- creditKNNRestDF[valid_instn,]
creditKNNTrainDF <- creditKNNRestDF[-valid_instn,]


# Training the KNN with k = 1
creditKNNTrainDF.X = creditKNNTrainDF[,c(2,3,4,6,7,10,12,17,18,20)]
creditKNNValidDF.X = creditKNNValidDF[,c(2,3,4,6,7,10,12,17,18,20)]
creditKNNTrainDF.Y = creditKNNTrainDF$PROFITABLE
creditKNNValidDF.Y = creditKNNValidDF$PROFITABLE
creditKNNTestDF.X = creditKNNTestDF[,c(2,3,4,6,7,10,12,17,18,20)]
creditKNNTestDF.Y = creditKNNTestDF$PROFITABLE

# Empty vectors to store accuracier
kValue = c()
validAcc = c()
trainAcc = c()

#For loop to iterate over various k values and then get the accuracies for both the #training set and Validation set
for(i in c(1, 3, 5, 7, 11, 15, 21, 25, 31, 35)){
  knnPredValid = knn(creditKNNTrainDF.X, creditKNNValidDF.X, creditKNNTrainDF.Y, k = i)
  knnPredTrain = knn(creditKNNTrainDF.X, creditKNNTrainDF.X, creditKNNTrainDF.Y, k = i)
  kValue = c(kValue, i)
  validAcc = c(validAcc, class_performance(table(creditKNNValidDF.Y, knnPredValid))[1])
  trainAcc = c(trainAcc, class_performance(table(creditKNNTrainDF.Y, knnPredTrain))[1])
}

knnDF = data.frame(kValue, validAcc, trainAcc)

```


2) Using the output, plot the accuracy for each value of k on both the training data and validation data

```{r}

ggplot() +
  geom_line(data = knnDF, aes(x = kValue , y = trainAcc), color = "blue") +
  geom_line(data = knnDF, aes(x = kValue, y = validAcc), color = 'red') +
  xlab('K values') +
  ylab('Accuracy') +
  ggtitle("Accuracy for Various K value")

```

The best value of K is 5, which has the highest accuracy on the validation set.

## Part 5: Comparing all the three models.

1) Calculate the accuracy of all three models on the testing data.

```{r}
AccuracyLogistic = class_performance(confusion_matrix(logisticTestPrediction,creditTestDF$PROFITABLE, 0.6619))[1]
print(AccuracyLogistic)

knnPredTest = knn(creditKNNTrainDF.X, creditKNNTestDF.X, creditKNNTrainDF.Y, k = 5)
class_performance(table(creditKNNTestDF.Y, knnPredTest))[1]

testAccuracy = predict_and_classify(prunedTree, creditTestDF, 
                                   creditTestDF$PROFITABLE, 0.5)
testAccuracy

```

Accuracy on the test data
Logistic Regression Accuracy = 68.33%
Tree Accuracy = 71%
KNN Accuracy = 70%

The Classification Trees are very good at predicting!
